---
title: "Impact of providing ART to Medicare ineligibles"
author: "Richard T. Gray"
-name: "Richard T. Gray"
affiliation: The Kirby Institute
date: Latest version - `r format(Sys.Date(), format="%B %d %Y")` <!---Add date automatically  using
  R whenever markdown is run! --->
output:
  word_document:
    reference_docx: mystyles.docx
bibliography: ../docs/medicare_ineligibles.bib    
---

This document summarises a simple methodolgy to calculate the impact of providing anti-retroviral therapy (ART) for free to people living with HIV (PLHIV) in Australia who are medicare ineligible. This analysis uses data from the ATRAS observational study [@petoumenos2013australian]

This document is written in dynamic format using R markdown v2 within R studio 0.98.1056 (using `r substr(R.Version()$version.string,1,15)`). Further details are available in the associated R markdown file which also contains the R code to produce all the results when the markdown is run. Code blocks have been supressed in the outputted document. <!--- The document is produced using R markdown within Rstudio with knitr, pandoc, Miktex producing the associated PDF output. To update the results the markdown file needs to be loaded in Rstudio or run using the tools separately. ---> 
<!----Testing Git Cloning --->

The R code for these calculations is available in the associated Rmarkdown file.

### Methodogy

#### Demographics

- summarise split in population
- growth and change in population over time

#### HIV transmission to partners

- summarise calculations for each year

#### Parameter table

- incoporate a full parmater table 

```{r,echo=FALSE}
# R code to define parameters for simulations
beta <- 0.02900653  
```

**Table 1** - Calculation input parameters 

Parameter  | Description | Value                                                        | Range | References
-----------|-------------|--------------------------------------------------------------|-------|------------
$\beta$    |             | `r signif(beta,2)`                                           |       |
$\mu$      |             | `r signif(beta,2)`                                           |       |
$\epsilon$ |             | `r signif(beta,2)`                                          |       |

### Results

```{r,echo = FALSE}
# R code to run the simulations
```

**Figure 1** - Change in new infections.

```{r, echo=FALSE}
# Do some pretty plots of the output
```

### Discussion


### References





