---
title: "Impact of providing ART to Medicare ineligibles"
author: "Richard T. Gray"
date: Latest version - `r format(Sys.Date(), format="%B %d %Y")` <!---Add date automatically  using R whenever markdown is run! --->
output: 
  word_document:
      reference_docx: mystyles.docx
---

This document summarises a simple methodolgy to calculate the impact of providing anti-retroviral therapy (ART) for free to people living with HIV (PLHIV) in Australia who are medicare ineligible. 

<!----Testing Git Cloning --->

The R code for these calculations is available in the associated Rmarkdown file.

### Methodogy

#### Demographics

- summarise split in population
- growth and change in population over time

#### HIV transmission to partners

- summarise calculations for each year

#### Parameter table

- incoporate a full parmater table 

```{r,echo=FALSE}
# R code to define parameters for simulations
beta <- 0.02900653  
```

**Table 1** - Calculation input parameters 

Parameter  | Description | Value                                                        | Range | References
-----------|-------------|--------------------------------------------------------------|-------|------------
$\beta$    |             | `r signif(beta,2)`                                           |       |
$\mu$      |             | `r signif(beta,2)`                                           |       |
$\epsilon$ |             | `r signif(beta,2)`                                          |       |

### Results

```{r,echo = FALSE}
# R code to run the simulations
```

**Figure 1** - Change in new infections.

```{r, echo=FALSE}
# Do some pretty plots of the output
```

### Discussion


### References





